---
title: "TS-Final Project-VAR"
output: html_document
---

```{r}
suppressWarnings(suppressMessages(library(fpp)))
suppressMessages(library(tseries))
suppressMessages(library(ggplot2))
suppressMessages(library(forecast))
suppressMessages(library(MuMIn))
suppressMessages(library(Metrics))
suppressMessages(library(MLmetrics))
suppressMessages(library(SciViews))
suppressMessages(library(TSA))
suppressMessages(library(vars))
suppressMessages(library(tidyverse))
suppressMessages(library(dplyr))
suppressMessages(library(prophet))
```


```{r}
Wanliu_data_daily <- read.csv(file = '/Users/amyzhang/Desktop/MScA/MSCA 31006/Project/Wanliu_data_daily.csv')
head(Wanliu_data_daily)
```

```{r}
PM2.5 <- Wanliu_data_daily$PM2.5
TEMP <- Wanliu_data_daily$TEMP
PRES <- Wanliu_data_daily$PRES
RAIN <- Wanliu_data_daily$RAIN
```

```{r}
# Plot the time series from 2013 March to 2014 December
daily_PM_2.5 <- ts(PM2.5, start = 2013.3, end = 2014.3, frequency = 365)
tsdisplay(daily_PM_2.5)
```

```{r}
# Create a daily Date object
inds <- seq(as.Date("2013-03-01"), as.Date("2017-02-28"), by = "day")
```

```{r}
ts_pm <- ts(PM2.5, start = c(2013, as.numeric(format(inds[1], "%j"))), frequency = 365)
ts_temp <- ts(TEMP, start = c(2013, as.numeric(format(inds[1], "%j"))), frequency = 365)
ts_pres <- ts(PRES, start = c(2013, as.numeric(format(inds[1], "%j"))), frequency = 365)
ts_rain <- ts(RAIN, start = c(2013, as.numeric(format(inds[1], "%j"))), frequency = 365)
```

```{r}
lambda_pm2.5 <- BoxCox.lambda(ts_pm)
lambda_pm2.5

lambda_temp <- BoxCox.lambda(ts_temp)
lambda_temp

lambda_pres <- BoxCox.lambda(ts_pres)
lambda_pres

lambda_pres <- BoxCox.lambda(ts_rain)
lambda_pres
```


```{r}
# Take first diff for pm 2.5 to get rid of non-stationary
pm_d1 <- diff(ts_pm, differences=1)
autoplot(ts_pm, main='PM 2.5')
autoplot(pm_d1, main='PM 2.5 (Differenced)')

temp_d1 <- diff(ts_temp, differences=1)
autoplot(temp_d1, main='Temp (Differenced)')

pres_d1 <- diff(ts_pres, differences=1)
autoplot(pres_d1, main='Pressure (Differenced)')

rain_d1 <- diff(ts_rain, differences=1)
autoplot(rain_d1, main='Pressure (Differenced)')
```

```{r}
# Box-Cox Transformation for temparature ==> Yes
ts.plot(ts_pm)
ts.plot(BoxCox(ts_pm, lambda_pm2.5))

qqnorm(ts_pm, main=expression(Normal~~Q-Q~~Plot))
qqline(ts_pm)

qqnorm(BoxCox(ts_pm, lambda_pm2.5),main=expression(Normal~~Q-Q~~Plot))
qqline(BoxCox(ts_pm, lambda_pm2.5))

hist(ts_pm)
hist(BoxCox(ts_pm, lambda_pm2.5))
```

```{r}
# Box-Cox Transformation for temparature ==> No, lambda is small, and normality decreased
autoplot(ts_temp)
autoplot(BoxCox(ts_temp, lambda_temp))

qqnorm(ts_temp, main=expression(Normal~~Q-Q~~Plot))
qqline(ts_temp)

qqnorm(BoxCox(ts_temp, lambda_temp),main=expression(Normal~~Q-Q~~Plot))
qqline(BoxCox(ts_temp, lambda_temp))
```

```{r}
# Box-Cox Transformation for pressure ==> No, didn't change much
autoplot(ts_pres)
autoplot(BoxCox(ts_pres, lambda_pres))

qqnorm(ts_pres, main=expression(Normal~~Q-Q~~Plot))
qqline(ts_pres)

qqnorm(BoxCox(ts_pres, lambda_pres),main=expression(Normal~~Q-Q~~Plot))
qqline(BoxCox(ts_pres, lambda_pres))

hist(ts_pres)
hist(BoxCox(ts_pres, lambda_pres))
```

```{r}
# Box-Cox Transformation for rain ==> No, it's mess
autoplot(ts_rain)

qqnorm(ts_rain, main=expression(Normal~~Q-Q~~Plot))
qqline(ts_rain)

hist(ts_rain)
```

```{r}
# Apply decompose function and visualize
fit_pm <- decompose(ts_pm, type="additive")
plot(fit_pm)

fit_temp <- decompose(ts_temp, type="additive")
plot(fit_temp)

fit_pres <- decompose(ts_pres, type="additive")
plot(fit_pres)

fit_train <- decompose(ts_rain, type="additive")
plot(fit_train)
```


```{r}
# Time Series plot of original data
tsdisplay(ts_pm)
tsdisplay(ts_temp)
tsdisplay(ts_pres)
tsdisplay(ts_rain)
```

```{r}
# Hypothesis Testing: ADF/KPSS Level/KPSS Trend
adf.test(ts_pm) # p-value < 0.05 ==> Reject NULL hypothesis ==> stationary
adf.test(ts_temp) # p-value > 0.05 ==> Acccept NULL hypothesis ==> non-stationary
adf.test(ts_pres) # p-value > 0.05 ==> Acccept NULL hypothesis ==> non-stationary

kpss.test(ts_pm) # p-value < 0.05 ==> Reject NULL hypothesis ==> non-stationary
kpss.test(ts_temp) # p-value < 0.05 ==> Reject NULL hypothesis ==> non-stationary
kpss.test(ts_pres) # p-value < 0.05 ==> Reject NULL hypothesis ==> non-stationary

kpss.test(ts_pm, null = "Trend") # p-value > 0.05 ==> Accept NULL hypothesis ==> stationary
kpss.test(ts_temp, null = "Trend") # p-value < 0.05 ==> Reject NULL hypothesis ==> non-stationary
kpss.test(ts_pres, null = "Trend") # p-value < 0.05 ==> Reject NULL hypothesis ==> non-stationary
```

```{r}
adf.test(ts_rain) # p-value < 0.05 ==> Reject NULL hypothesis ==> stationary
kpss.test(ts_rain) # p-value > 0.05 ==> Accept NULL hypothesis non-stationary
kpss.test(ts_rain, null = "Trend") # p-value > 0.05 ==> Accept NULL hypothesis ==> stationary
```

# VAR model
```{r}
x = cbind(ts_pm, ts_temp, ts_pres)
plot.ts(x, main = "PM 2.5, Temperature, and Pressure (Logged)", xlab = "Time")
```

```{r}
x = cbind(ts_pm, ts_temp, ts_pres)
VARselect(x, lag.max = 10, type = 'both')$selection
```

# The R output shows the lag length selected by each of the information criteria available in the vars package. There is a large discrepancy between the VAR(5) selected by the AIC and the VAR(3) selected by the BIC.

# As a result, we first fit a VAR(5), as selected by the AIC, and VAR(3) selected by the BIC.
# For each model we test that the residuals are uncorrelated using a Portmanteau test.
# The null hypothesis of no serial correlation in the residuals is rejected for both VAR(5) and a VAR(3), so we continue to VAR(6)

```{r}
#x_df = cbind(ts_pm, ts_temp, ts_pres, ts_rain)
fit1 <- VAR(x, p=5, type="both")
serial.test(fit1, lags.pt = 10, type="PT.asymptotic")
fit2 <- VAR(x, p=3, type="both")
serial.test(fit2, lags.pt = 10, type="PT.asymptotic")
summary(fit2)
```

```{r}
var6 <- VAR(x, p=6, type="both")
serial.test(var6, lags.pt = 10, type="PT.asymptotic")
```

```{r}
summary(var6)
```

```{r}
pm_d1 <- diff(PM2.5, differences=1)
temp_d1 <- diff(TEMP, differences=1)
pres_d1 <- diff(PRES, differences=1)
x_df <- cbind(pm_d1, temp_d1, pres_d1)
var9 <- VAR(x_df, p=10, type="both")
VARselect(x_df, lag.max = 10, type = "both")$selection
serial.test(var9, lags.pt = 10, type="PT.asymptotic")
#summary(var9)
```

```{r}
forecast(var6) %>%
  autoplot() + xlab("Year")
```

```{r}
f <- forecast(var6, h=30) %>% autoplot() + coord_cartesian(xlim = c(2016.1,2017.2438))
f
```

## Ignore the below since we're only for differenced data ###
```{r}
# Predict
# pred1 <- predict(fit1, n.ahead = 30)
# pred2 <- predict(fit2, n.ahead = 30)

# Inverse difference prediction for PM 2.5
# pred1_inv_pm <- diffinv(pred1$fcst$pm_d1[, 1], xi = ts_pm[seq_len(1)])
# pred2_inv_pm <- diffinv(pred2$fcst$pm_d1[, 1], xi = ts_pm[seq_len(1)])

# Inverse difference prediction for temperature
# pred1_inv_tp <- diffinv(pred1$fcst$temp_d1[, 1], xi = mean(ts_temp))
# pred2_inv_tp <- diffinv(pred2$fcst$temp_d1[, 1], xi = ts_temp[seq_len(1)])

# Inverse difference prediction for pressure
# pred1_inv_ps <- diffinv(pred1$fcst$pres_d1[, 1], xi = ts_pres[seq_len(1)])
# pred2_inv_ps <- diffinv(pred2$fcst$pres_d1[, 1], xi = ts_pres[seq_len(1)])

# Do not need to inverse diff for rain
# pred1_inv_rn <- pred1$fcst$ts_rain_2[, 1]
# pred2_inv_rn <- pred2$fcst$ts_rain_2[, 1]

# pred1_inv_pm_ts <- ts(pred1_inv_pm, start = c(2017, as.numeric(format(as.Date("2017-03-30"), "%j"))), frequency = 365)
# ts.plot(cbind(window(ts_pm), pred1_inv_pm_ts), lty = 1:2)
# title("PM2.5 forecast with VAR(9)")

# pred1_inv_tp_ts <- ts(pred1_inv_tp, start = c(2017, as.numeric(format(as.Date("2017-03-30"), "%j"))), frequency = 365)
# ts.plot(cbind(window(ts_temp), pred1_inv_tp_ts), lty = 1:2)
# title("PM2.5 forecast with VAR(9)")

# pred1_inv_ps_ts <- ts(pred1_inv_ps, start = c(2017, as.numeric(format(as.Date("2017-03-30"), "%j"))), frequency = 365)
# ts.plot(cbind(window(ts_pres), pred1_inv_ps_ts), lty = 1:2)
# title("PM2.5 forecast with VAR(9)")

# pred1_inv_ps_rn <- ts(pred1_inv_rn, start = c(2017, as.numeric(format(as.Date("2017-03-30"), "%j"))), frequency = 365)
# ts.plot(cbind(window(ts_rain_2), pred1_inv_ps_rn), lty = 1:2)
# title("PM2.5 forecast with VAR(9)")

###Plots for VAR(3) model###
# pred2_inv_pm_ts <- ts(pred2_inv_pm, start = c(2017, as.numeric(format(as.Date("2017-03-30"), "%j"))), frequency = 365)
# ts.plot(cbind(window(ts_pm), pred2_inv_pm_ts), lty = 1:2)
# title("PM2.5 forecast with VAR(3)")

# pred2_inv_tp_ts <- ts(pred2_inv_tp, start = c(2017, as.numeric(format(as.Date("2017-03-30"), "%j"))), frequency = 365)
# ts.plot(cbind(window(ts_temp), pred2_inv_tp_ts), lty = 1:2)
# title("PM2.5 forecast with VAR(3)")

# pred2_inv_ps_ts <- ts(pred2_inv_ps, start = c(2017, as.numeric(format(as.Date("2017-03-30"), "%j"))), frequency = 365)
# ts.plot(cbind(window(ts_pres), pred2_inv_ps_ts), lty = 1:2)
# title("PM2.5 forecast with VAR(3)")

# pred2_inv_ps_rn <- ts(pred2_inv_rn, start = c(2017, as.numeric(format(as.Date("2017-03-30"), "%j"))), frequency = 365)
# ts.plot(cbind(window(ts_rain_2), pred2_inv_ps_rn), lty = 1:2)
# title("PM2.5 forecast with VAR(3)")
```


```{r}
# Time Series Plot of residuals of VAR(9)
ts.plot(resid(fit3)[, 1])
ts.plot(resid(fit3)[, 2])
ts.plot(resid(fit3)[, 3])

# ACF Plot of residuals of VAR(9)
acf(resid(fit3)[, 1])
acf(resid(fit3)[, 2])
acf(resid(fit3)[, 3])

# PACF Plot of residuals of VAR(9)
pacf(resid(fit3)[, 1])
pacf(resid(fit3)[, 2])
pacf(resid(fit3)[, 3])

# Histogram of residuals of VAR(9)
hist(resid(fit3)[, 1])
hist(resid(fit3)[, 2])
hist(resid(fit3)[, 3])
```


### CV begins here ###
```{r}
data <- read.csv(file = '/Users/amyzhang/Desktop/MScA/MSCA 31006/Project/Wanliu_data_daily.csv')
#firstHour <- 24*(as.Date("2013-03-01 00:00:00")-as.Date("2013-1-1 00:00:00"))
data_ts <- ts(data=data[c('PM2.5', 'PRES', 'TEMP', 'RAIN')], start=c(2013,3), frequency=365.25)

PM2.5 <- data_ts[, 'PM2.5']
TEMP <- data_ts[, 'TEMP']
PRES <- data_ts[, 'PRES']

x = cbind(PM2.5, TEMP, PRES)

# CV
k=1381
h=30
p=365.25
n <- length(PM2.5)

st <- tsp(PM2.5)[1]+(k-2)/p #  gives the start time in time units,

error_1_var <- matrix(NA,n-k,h)
error_2_var <- matrix(NA,n-k,h)

AIC_1_var <- matrix(NA,n-k)
AIC_2_var <- matrix(NA,n-k)

i<-1

for(i in (1:80))
{
  ### One Month rolling forecasting
  # Expanding Window 
  train_1 <- window(x, end=st + i/p)  ## Window Length: k+i
  
  # Sliding Window - keep the training window of fixed length. 
  # The training set always consists of k observations.
  train_2 <- window(x, start=st+(i-k+1)/p, end=st+i/p) ## Window Length: k
  test <- window(PM2.5, start=st + (i+1)/p, end=st + (i+h)/p) ## Window Length: H
  
  if (i<4) {
    cat(c("*** CV", i,":","len(Expanding Window):",length(train_1), "len(Sliding Window):",length(train_2), "len(Test):",length(test),'\n'  ))
    cat(c("*** TRAIN -  Expanding WIndow:",tsp(train_1)[1],'-',tsp(train_1)[2],'\n'))
    cat(c("*** TRAIN - Sliding WIndow:",tsp(train_2)[1],'-',tsp(train_2)[2],'\n'))
    #cat(c("*** TEST:",tsp(test)[1],'-',tsp(test)[2],'\n'))
    cat("*************************** \n \n")
  }

  # M1 - VAR(6) with expanding window
  fit_1 <- VAR(train_1, p=6, type="both")
  fcast_1 <- forecast(fit_1, h=h)
  fcast_1_p <- predict(fit_1, n.ahead=h)
  
  mean_1 <- fcast_1_p$fcst$PM2.5[,1]
  #fcast_1_ts <- ts(mean_1, start = c(2020, 242), end = c(2020, 271), frequency = 365.25)
  
  s1 <- summary(fit_1)
  
  # M2 -VAR(6) with sliding window
  fit_2 = VAR(train_2, p=6, type='both')
  fcast_2 <- forecast(fit_2, h=h)
  fcast_2_p <- predict(fit_2, n.ahead=h)
  
  mean_2 <- fcast_2_p$fcst$PM2.5[,1]
  #fcast_2_ts <- ts(mean_2, start = c(2020, 242), end = c(2020, 271), frequency = 365.25)
  
  s2 <- summary(fit_2)
  error_1_var[i,1:length(test)] <- mean_1[1:length(test)] - test
  
  error_2_var[i,1:length(test)] <- mean_2[1:length(test)] - test
  
  AIC_1_var[i, 1] <- ln(det(s1$covres)) + (2*p*k^2)/n
  AIC_2_var[i, 1] <- ln(det(s2$covres)) + (2*p*k^2)/n
  i<-i+1
}

```

#[1]  73.625000  39.416667  93.500000 162.541667 236.666667  13.250000  14.291667  83.541667  33.125000
[10]   5.437500   5.375000  45.583333  99.750000  78.708333 187.208333 241.250000 117.354167  21.333333
[19]  87.333333  97.416667  10.479167  58.020833  72.333333  15.916667  21.458333   8.250000  27.125000
[28]  64.000000   9.958333

 
